---
---
title: "Analysis of the number of people living in a household"
author: "Group 5"
output:
  pdf_document:
    latex_engine: xelatex
    number_sections: no
  html_document:
    df_print: paged
fig_caption: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, eval = TRUE, warning = FALSE, message = FALSE, comment = NA)
```

```{r libraries}
library(tidyverse)
library(moderndive)
library(skimr)
library(kableExtra)
library(gridExtra)
library(olsrr)
library(broom)
```


# Introduction {#sec:Intro}
Experiments were conducted as part of research into "The number of people living in a household".

1788 households were used in the experiments, and they each had their Annual household income in Philippine peso(Total.Household.Income), the region of the Philippines(Region), annual expenditure by the household on food in Philippine peso(Total.Food.Expenditure), head of the households sex(Household.Head.Sex), head of the households age in years(Household.Head.Age), relationship between the group of people living in the house(Type.of.Household), number of people living in the house(Total.Number.of.Family.members), floor area of the house(House.Floor.Area) in $m^2$, Age of the building in years(House.Age), number of bedrooms in the house(Number.of.bedrooms), whether the house have electricity with 1 is Yes and 0 is No)(Electricity).

No missing or undefined values are included in the dataset, and Total.Number.of.Family.members is given against the attributes of each household, including Total.Household.Income, Region, Total.Food.Expenditure, Household.Head.Sex, Household.Head.Age, Type.of.Household, House.Floor.Area, House.Age, Number.of.bedrooms and Electricity.

This dataset is helpful in a simple but illuminating research in understanding which household related variables may affect the number of people living in a household.

In particular, this report presents numerical and graphical summaries of the number of people living in a household, and shows which variables have an effect on the it. Then, a linear model was fitted to estimate the difference, on average, between the number of people living in a household of household head male and female, and among the number of people living in a household of three types of household.

```{r loading dataset}
data <- read.csv('dataset5.csv') # Loading data set
```

```{r data cleaning}
data$Number.of.bedrooms<-as.factor(data$Number.of.bedrooms)
data$Electricity<-as.factor(data$Electricity)
```

# Exploratory Data Analysis {#sec:EDA}
```{r}
# Summarize the Household.Head.Sex variable
data %>% group_by(Household.Head.Sex) %>%
  summarise(n=n(),Mean=round(mean(Total.Number.of.Family.members),digits=1), St.Dev=round(sd(Total.Number.of.Family.members),digits=1), # Summarize info
  Min=min(Total.Number.of.Family.members), Q1 = quantile(Total.Number.of.Family.members,0.25), Median=median(Total.Number.of.Family.members),
  Q3 = quantile(Total.Number.of.Family.members,0.75), Max=max(Total.Number.of.Family.members))%>%
  kbl(booktabs = T,caption = '\\label{tab:summarise1}Summary statistics on Total.Number.of.Family.members by the Household.Head.Sex')%>%
  kable_styling(latex_options = c("striped","hold_position"))
```

Table \ref{tab:summarise1} shows that there are roughly fourth as many male household head in the sample (`r nrow(data[data$Household.Head.Sex=="Male",])` compared to `r nrow(data[data$Household.Head.Sex=="Female",])`). 
The summaries of the total number of family members of male household head were roughly similar with the corresponding summaries of female household head except mean. The mean of the total family members of male household head was `r round(mean(data[data$Household.Head.Sex=="Male",]$Total.Number.of.Family.members),1)` compared to `r round(mean(data[data$Household.Head.Sex=="Female",]$Total.Number.of.Family.members),1)` for female household head.

\newpage
```{r}
# Summarize the Type.of.Household variable
data %>% group_by(Type.of.Household) %>%
  summarise(n=n(),Mean=round(mean(Total.Number.of.Family.members),digits=1), St.Dev=round(sd(Total.Number.of.Family.members),digits=1), # Summarize info
  Min=min(Total.Number.of.Family.members), Q1 = quantile(Total.Number.of.Family.members,0.25), Median=median(Total.Number.of.Family.members),
  Q3 = quantile(Total.Number.of.Family.members,0.75), Max=max(Total.Number.of.Family.members))%>%
  kbl(booktabs = T,caption = '\\label{tab:summarise2}Summary statistics on Total.Number.of.Family.members by the Type.of.Household')%>%
  kable_styling(latex_options = c("striped","hold_position"))
```

Table \ref{tab:summarise2} shows that there are significant difference in terms of the number of different household types in the sample, with (`r nrow(data[data$Type.of.Household=="Extended Family",])` extended family, `r nrow(data[data$Type.of.Household=="Single Family",])` single family and `r nrow(data[data$Type.of.Household=="Two or More Nonrelated Persons/Members",])` two or more nonrelated persons/members). 
The summaries of the total number of family members of single family were slightly lower than the corresponding summaries of other household types. For example, the means of the total family members of extended family, single family and two or more nonrelated persons/members were `r round(mean(data[data$Type.of.Household=="Extended Family",]$Total.Number.of.Family.members),1)`,`r round(mean(data[data$Type.of.Household=="Single Family",]$Total.Number.of.Family.members),1)` and `r round(mean(data[data$Type.of.Household=="Two or More Nonrelated Persons/Members",]$Total.Number.of.Family.members),1)` respectively.

```{r}
# Summarize the Number.of.bedrooms variable
data %>% group_by(Number.of.bedrooms) %>%
  summarise(n=n(),Mean=round(mean(Total.Number.of.Family.members),digits=1), St.Dev=round(sd(Total.Number.of.Family.members),digits=1), # Summarize info
  Min=min(Total.Number.of.Family.members), Q1 = quantile(Total.Number.of.Family.members,0.25), Median=median(Total.Number.of.Family.members),
  Q3 = quantile(Total.Number.of.Family.members,0.75), Max=max(Total.Number.of.Family.members))%>%
  kbl(booktabs = T,caption = '\\label{tab:summarise3}Summary statistics on Total.Number.of.Family.members by the Number.of.bedrooms')%>%
  kable_styling(latex_options = c("striped","hold_position"))
```

Table \ref{tab:summarise3} shows that there are significant difference in terms of the number of bedrooms in the sample, with (`r nrow(data[data$Number.of.bedrooms==0,])` zero bedroom, `r nrow(data[data$Number.of.bedrooms==1,])` one bedroom, `r nrow(data[data$Number.of.bedrooms==2,])` two bedroomds, `r nrow(data[data$Number.of.bedrooms==3,])` three bedrooms, `r nrow(data[data$Number.of.bedrooms==4,])` four bedrooms, `r nrow(data[data$Number.of.bedrooms==5,])` five bedrooms and `r nrow(data[data$Number.of.bedrooms==6,])` six bedrooms). The summaries of the total number of family members of different number of bedrooms were different, mainly in terms of mean and max.

```{r}
# Summarize the Electricity variable
data %>% group_by(Electricity) %>%
  summarise(n=n(),Mean=round(mean(Total.Number.of.Family.members),digits=1), St.Dev=round(sd(Total.Number.of.Family.members),digits=1), # Summarize info
  Min=min(Total.Number.of.Family.members), Q1 = quantile(Total.Number.of.Family.members,0.25), Median=median(Total.Number.of.Family.members),
  Q3 = quantile(Total.Number.of.Family.members,0.75), Max=max(Total.Number.of.Family.members))%>%
  kbl(booktabs = T,caption = '\\label{tab:summarise4}Summary statistics on Total.Number.of.Family.members by the Electricity')%>%
  kable_styling(latex_options = c("striped","hold_position"))
```

Table \ref{tab:summarise4} shows that there are roughly fourth as many households with electricity in the sample (`r nrow(data[data$Electricity==1,])` compared to `r nrow(data[data$Electricity==0,])`). 
The summaries of the total number of family members of households with electricity were roughly similar with the corresponding summaries of households without electricity except mean and max. For example, the mean of the total family members of households without electricity was `r round(mean(data[data$Electricity==0,]$Total.Number.of.Family.members),1)` compared to `r round(mean(data[data$Electricity==1,]$Total.Number.of.Family.members),1)` for households with electricity.

```{r boxplot, fig.align = "center", fig.cap = "\\label{fig:box} Level.of.development and European.Union.Membership by Total.Number.of.Family.members respectively.", fig.pos = 'H'}
# Draw boxplots
box1 <- ggplot(data,aes(x=Household.Head.Sex,y=Total.Number.of.Family.members))+
  geom_boxplot()+
  labs(x='Household.Head.Sext',y='Women.Entrepreneurship.Index')
box2 <- ggplot(data,aes(x=Type.of.Household,y=Total.Number.of.Family.members))+
  geom_boxplot()+
  labs(x='Type.of.Household',y='Total.Number.of.Family.members')
box3 <- ggplot(data,aes(x=Number.of.bedrooms,y=Total.Number.of.Family.members))+
  geom_boxplot()+
  labs(x='Number.of.bedrooms',y='Total.Number.of.Family.members')
box4 <- ggplot(data,aes(x=Electricity,y=Total.Number.of.Family.members))+
  geom_boxplot()+
  labs(x='Electricity',y='Total.Number.of.Family.members')

# Merge two boxplots
grid.arrange(box1,box2,box3,box4,nrow=2)
```
The top boxplot shows that female household head and single family household type have lower total number of family members.
Thus, the household head sex and type of household have significantly effect on the total number of family members. 

```{r plot1,echo = FALSE, fig.align = "center",fig.cap = "\\label{fig:plot1}Scatterplots of Total.Number.of.Family.members against total household income, total food expenditure, household head age, house floor area and house age respectively", fig.pos = 'H'}

# Draw scatter plots
point1 <-ggplot(data,aes(x=Total.Household.Income,y=Total.Number.of.Family.members))+
  geom_point()+
  labs(x='Total.Household.Income',y='Total.Number.of.Family.members')+
  geom_smooth(method = "lm", se = FALSE)

point2 <- ggplot(data,aes(x=Total.Food.Expenditure,y=Total.Number.of.Family.members))+
  geom_point()+
  labs(x='Total.Food.Expenditure',y='Total.Number.of.Family.members')+
  geom_smooth(method = "lm", se = FALSE)

point3 <- ggplot(data,aes(x=Household.Head.Age,y=Total.Number.of.Family.members))+
  geom_point()+
  labs(x='Household.Head.Age',y='Total.Number.of.Family.members')+
  geom_smooth(method = "lm", se = FALSE)

point4 <- ggplot(data,aes(x=House.Floor.Area,y=Total.Number.of.Family.members))+
  geom_point()+
  labs(x='House.Floor.Area',y='Total.Number.of.Family.members')+
  geom_smooth(method = "lm", se = FALSE)

point5 <- ggplot(data,aes(x=House.Age,y=Total.Number.of.Family.members))+
  geom_point()+
  labs(x='House.Age',y='Total.Number.of.Family.members')+
  geom_smooth(method = "lm", se = FALSE)

# Merge two scatter plot
grid.arrange(point1,point2,point3,point4,point5,nrow=3)
```


Figure \ref{fig:plot1} displays scatterplots of total number of family members against total household income, total food expenditure, household head age, house floor area and house age respectively. Since there is one ourlier shown in these scatter plots, the outlier is excluded from the dataset and new scatterplots are displayed in the following.

```{r plot2,echo = FALSE, fig.align = "center",fig.cap = "\\label{fig:plot2}Scatterplots of Total.Number.of.Family.members against total household income, total food expenditure, household head age, house floor area and house age respectively", fig.pos = 'H'}

#removing outlier
data<-subset(data,House.Floor.Area<=600)
data<-subset(data,Total.Household.Income <6071029)

# Draw scatter plots
point1 <-ggplot(data,aes(x=Total.Household.Income,y=Total.Number.of.Family.members))+
  geom_point()+
  labs(x='Total.Household.Income',y='Total.Number.of.Family.members')+
  geom_smooth(method = "lm", se = FALSE)

point2 <- ggplot(data,aes(x=Total.Food.Expenditure,y=Total.Number.of.Family.members))+
  geom_point()+
  labs(x='Total.Food.Expenditure',y='Total.Number.of.Family.members')+
  geom_smooth(method = "lm", se = FALSE)

point3 <- ggplot(data,aes(x=Household.Head.Age,y=Total.Number.of.Family.members))+
  geom_point()+
  labs(x='Household.Head.Age',y='Total.Number.of.Family.members')+
  geom_smooth(method = "lm", se = FALSE)

point4 <- ggplot(data,aes(x=House.Floor.Area,y=Total.Number.of.Family.members))+
  geom_point()+
  labs(x='House.Floor.Area',y='Total.Number.of.Family.members')+
  geom_smooth(method = "lm", se = FALSE)

point5 <- ggplot(data,aes(x=House.Age,y=Total.Number.of.Family.members))+
  geom_point()+
  labs(x='House.Age',y='Total.Number.of.Family.members')+
  geom_smooth(method = "lm", se = FALSE)

grid.arrange(point1,point2,point3,point4,point5,nrow=3)
```


```{r, eval = TRUE}
Cor.data.THI <-data %>%
  get_correlation(formula = Total.Number.of.Family.members ~  Total.Household.Income)

Cor.data.TFE <-data %>%
  get_correlation(formula = Total.Number.of.Family.members ~ Total.Food.Expenditure)

Cor.data.HHA <-data %>%
  get_correlation(formula = Total.Number.of.Family.members ~ Household.Head.Age)

Cor.data.HFA <-data %>%
  get_correlation(formula = Total.Number.of.Family.members ~ House.Floor.Area)

Cor.data.HA <-data %>%
  get_correlation(formula = Total.Number.of.Family.members ~ House.Age)

```
Figure \ref{fig:plot2} displays scatterplots of total number of family members against total household income, total food expenditure, household head age, house floor area and house age respectively, with corresponding correlation: `r round(Cor.data.THI, 3)`, `r round(Cor.data.TFE, 3)`,`r round(Cor.data.HHA, 3)`,`r round(Cor.data.HFA, 3)`,`r round(Cor.data.HA, 3)`.
All except household head age and house age have positive relationship with total number of family members.


# Formal Data Analysis {#sec:FDA}
This report fits the following generalized linear model for the research.


$$            \begin{array}{ll}
              \widehat{TNOFM}_{i} = \hat{a}+\hat{b} \cdot THI + \hat c \cdot TFE +\hat{d}_{Male} \cdot \mathbb{I}_{Male}(i) + \hat{e} \cdot HHA+\hat{f}_{SingleFamily} \cdot \mathbb{I}_{SingleFamily}(i) \\
              + \hat{g}_{Two Or More Nonrelated Persons/Members} \cdot \mathbb{I}_{Two Or More Nonrelated Persons/Members}(i) +\hat{h} \cdot HFA +\hat{i} \cdot HA \\
              + \hat{j}_{N1} \cdot \mathbb{I}_{N1}(i) +\hat{k}_{N2} \cdot \mathbb{I}_{N2}(i)+\hat{l}_{N3} \cdot \mathbb{I}_{N3}(i)+\hat{m}_{N4} \cdot \mathbb{I}_{N4}(i)+\hat{n}_{N5} \cdot \mathbb{I}_{N5}(i)+\hat{o}_{N6} \cdot \mathbb{I}_{N6}(i)+ \hat{p}_{Yes} \cdot \mathbb{I}_{Yes}(i)
            \end{array}$$
where

$\widehat{TNOFM}_{i}$ is the expected value of the total number of family members of $i^{th}$ observation

$\hat a$ is the intercept of the generalized linear model

$\hat b$ is the coefficient for the first explanatory variable "total household income"

$\hat c$ is the coefficient for the second explanatory variable "total food expenditure"

$\hat d$ is the coefficient for the third explanatory variable "household head sex"

$\mathbb{I}_{Male}(x)$ is an indicator function:
$$\mathbb{I}_{Male}(x)=\left\{
            \begin{array}{ll}
              1 ~~~ \mbox{if the observation is male} ~ \mbox{},\\
              0 ~~~ \mbox{Otherwise}.\\
            \end{array}
          \right.$$

$\hat e$ is the coefficient for the explanatory variable "household head age"

$\hat f$ and $\hat g$ are the coefficient of the classification variable "type of household"

$\mathbb{I}_{Single Family }(x)$ is an indicator function:
$$\mathbb{I}_{Single Family }(x)=\left\{
            \begin{array}{ll}
              1 ~~~ \mbox{if the observation is single family } ~ \mbox{},\\
              0 ~~~ \mbox{Otherwise}.\\
            \end{array}
          \right.$$

$\mathbb{I}_{Two Or More Nonrelated Persons/Members }(x)$ is an indicator function:
$$\mathbb{I}_{Two Or More Nonrelated Persons/Members }(x)=\left\{
            \begin{array}{ll}
              1 ~~~ \mbox{if the observation has two or more nonrelated persons/members } ~ \mbox{},\\
              0 ~~~ \mbox{Otherwise}.\\
            \end{array}
          \right.$$


$\hat h$ is the coefficient of the explanatory variable "house floor area"

$\hat i$ is the coefficient of the explanatory variable "house age"

$\hat j$,$\hat k$, $\hat l$, $\hat m$, $\hat n$, and $\hat o$ are the coefficient of classification variable "number of bedrooms"

$\mathbb{I}_{N1 }(x)$ is an indicator function:
$$\mathbb{I}_{N1 }(x)=\left\{
            \begin{array}{ll}
              1 ~~~ \mbox{if the house has one bedroom} ~ \mbox{},\\
              0 ~~~ \mbox{Otherwise}.\\
            \end{array}
          \right.$$
          
$\mathbb{I}_{N2 }(x)$ is an indicator function:
$$\mathbb{I}_{N2 }(x)=\left\{
            \begin{array}{ll}
              1 ~~~ \mbox{if the house has two bedrooms} ~ \mbox{},\\
              0 ~~~ \mbox{Otherwise}.\\
            \end{array}
          \right.$$

$\mathbb{I}_{N3 }(x)$ is an indicator function:
$$\mathbb{I}_{N3 }(x)=\left\{
            \begin{array}{ll}
              1 ~~~ \mbox{if the house has three bedrooms} ~ \mbox{},\\
              0 ~~~ \mbox{Otherwise}.\\
            \end{array}
          \right.$$

$\mathbb{I}_{N4 }(x)$ is an indicator function:
$$\mathbb{I}_{N4 }(x)=\left\{
            \begin{array}{ll}
              1 ~~~ \mbox{if the house has four bedrooms} ~ \mbox{},\\
              0 ~~~ \mbox{Otherwise}.\\
            \end{array}
          \right.$$

$\mathbb{I}_{N5 }(x)$ is an indicator function:
$$\mathbb{I}_{N5}(x)=\left\{
            \begin{array}{ll}
              1 ~~~ \mbox{if the house has five bedrooms} ~ \mbox{},\\
              0 ~~~ \mbox{Otherwise}.\\
            \end{array}
          \right.$$

$\mathbb{I}_{N6 }(x)$ is an indicator function:
$$\mathbb{I}_{N6 }(x)=\left\{
            \begin{array}{ll}
              1 ~~~ \mbox{if the house have six bedrooms} ~ \mbox{},\\
              0 ~~~ \mbox{Otherwise}.\\
            \end{array}
          \right.$$


$\hat p$ is the coefficient of the explanatory variable "electricity"

$\mathbb{I}_{Yes }(x)$ is an indicator function:
$$\mathbb{I}_{Yes }(x)=\left\{
            \begin{array}{ll}
              1 ~~~ \mbox{if the house have electricity} ~ \mbox{},\\
              0 ~~~ \mbox{Otherwise}.\\
            \end{array}
          \right.$$
\newpage
```{r}
full.model <- glm(Total.Number.of.Family.members~.,data = data,family = poisson)


summary(full.model)$coefficients%>%
  kbl(booktabs = T,caption = '\\label{tab:glm1} Estimates of the parameters from the Generalized linear model.',col.names = c('Estimate', "Std.Error", "Z value",'P value'))%>%
  kable_styling(latex_options = c("striped","hold_position"))


```

Table\ref{tab:glm1} shows that all of the classification variables number of bedroom is not significant due to all p value are greater than 0.05. The p value of electricity and house floor area are approximately 0.685 and 0.255 which are both greater than 0.05, so these variables are also not significant for the response variable. One of the categorical variables for type of household is not significant (two or more unrelated people in the household). The remaining variables are all significant. Due to some variable is not significant for response variable, this report use step function to select variable that contribute to the model.


```{r}
s <-step(full.model,trace=0)

summary(s)$coefficients%>%
  kbl(booktabs = T,caption = '\\label{tab:glm2} Estimates of the parameters from the Generalized linear model.',col.names = c('Estimate', "Std.Error", "Z value",'P value'))%>%
  kable_styling(latex_options = c("striped","hold_position"))

```

Table\ref{tab:glm2} shows the best model with smallest AIC.
$$            \begin{array}{ll}
              \widehat{TNOFM}_{i} = \hat{a}+\hat{b} \cdot THI + \hat c \cdot TFE +\hat{d}_{Male} \cdot \mathbb{I}_{Male}(i) + \hat{e} \cdot HHA+\hat{f}_{SingleFamily} \cdot \mathbb{I}_{SingleFamily}(i) \\
              + \hat{g}_{Two Or More Nonrelated Persons/Members} \cdot \mathbb{I}_{Two Or More Nonrelated Persons/Members}(i) +\hat{h} \cdot HA\\
            \end{array}$$
where

$\widehat{TNOFM}_{i}$ is the expected value of the total number of family members of $i^{th}$ observation

$\hat a$ is the intercept of the generalized linear model

$\hat b$ is the coefficient for the first explanatory variable "total household income"

$\hat c$ is the coefficient for the second explanatory variable "total food expenditure"

$\hat d$ is the coefficient for the third explanatory variable "household head sex"

$\mathbb{I}_{Male}(x)$ is an indicator function:
$$\mathbb{I}_{Male}(x)=\left\{
            \begin{array}{ll}
              1 ~~~ \mbox{if the observation is male} ~ \mbox{},\\
              0 ~~~ \mbox{Otherwise}.\\
            \end{array}
          \right.$$

$\hat e$ is the coefficient for the explanatory variable "household head age"

$\hat f$ and $\hat g$ are the coefficient of the classification variable "type of household"

$\mathbb{I}_{Single Family }(x)$ is an indicator function:
$$\mathbb{I}_{Single Family }(x)=\left\{
            \begin{array}{ll}
              1 ~~~ \mbox{if the observation is single family } ~ \mbox{},\\
              0 ~~~ \mbox{Otherwise}.\\
            \end{array}
          \right.$$

$\mathbb{I}_{Two Or More Nonrelated Persons/Members }(x)$ is an indicator function:
$$\mathbb{I}_{Two Or More Nonrelated Persons/Members }(x)=\left\{
            \begin{array}{ll}
              1 ~~~ \mbox{if the observation has two or more nonrelated persons/members } ~ \mbox{},\\
              0 ~~~ \mbox{Otherwise}.\\
            \end{array}
          \right.$$
$\hat h$ is the coefficient of the explanatory variable "house age"
